\chapter{Introduction}

\section{Preamble}

The {\tt otagothesis} package consists a style file and document
template for writing theses and dissertations at the University of
Otago.  It was intended only for computer science students, but seems
to have found wider popularity.  Its purpose is to take the hard work
of constructing a large document out of your hands.

Earlier versions of this package used Latex to produce a DVI file
which could be converted to PostScript and printed.  There are several
advantages to using Pdflatex instead: the document goes straight to
PDF, and may include graphics files of many types instead of only
EPSF.  This package now uses Pdflatex instead of Latex.  If you wish
to use it in place of an earlier version, you may need to convert any
EPSF diagrams to PDF with the \verb|epstopdf| program, and you should
remove the \verb|[dvips]| option from the \verb|\usepackage{graphicx}|
command.

This document is a quick tour of the {\tt otagothesis} package and a
very brief comment on how to organise a thesis in Computer Science.
For an even more brief introduction to the package, read the README
file.  Please bear in mind that this is not a style guide, nor an
example of good content---although you will find some ideas on how
to cite other authors' work, and how to include tables and figures in
your document.  These are all specific to the Systems Research Lab at
the Computer Science Department; all other students should seek advice
from their supervisors before following {\em any} of the advice
contained herein!

It should be very easy to convert the file {\tt thesis.tex} to
your own use.
The {\tt otagothesis.sty} file is not quite as flexible, but feel
free to hack around with it if you wish---just be sure to pass on
the original to others (or take my name off the changed version).
The idea is that if you are an Otago University student
you shouldn't need to change {\tt otagothesis.sty} at all.

\section{Files You Need}
\label{sec:files}
There are several files you need to have a copy of before you will be
able to compile this document for yourself. These files are included in the original set of files. They are:

\begin{itemize}
\item {\tt otagothesis.sty}: this is the file which provides the
{\tt otagothesis} package.  It contains some useful commands and
options, and formats your pages according to University library restrictions.
\item {\tt logo.pdf}: the Otago University crest, required for your
library declaration page.
\item {\tt natbib.sty}: this is the file which provides Harvard style
(author, date) citations.  It is provided in TeTeX, the distribution
of LaTeX that comes with Redhat Linux.
\item {\tt plainnat.bst}: this provides Author-date style bibliographies.
It is also standard in TeTeX.
\end{itemize}

The easiest way to use the otagothesis package is to put {\tt
otagothesis.sty} and {\tt logo.pdf} in the same directory as your
document.  At its simplest, put \verb|\usepackage{otagothesis}|
just after the \verb|\documentclass{report}| declaration, and you're
away.  For other usage options, follow the pattern in the
\verb|thesis.tex| file.

\section{Compiling}
Just to see if everything is working, try compiling this document by
unzipping the distribution, changing into the \verb|example_document|
directory, and
typing `make'.  You should see a whole lot of messages printed to the
screen, followed by a preview window showing you a title-page.  If you
don't get this, try reading Subsection~\ref{subsec:trouble} on
troubleshooting the distribution.  If everything seems to be OK, go on
to Section~\ref{sec:overview} which tells you a bit about using these files.

If you don't like previewing your document every time you recompile
it, read Subsection~\ref{subsec:makefile} which has some suggestions
for changing the makefile.

\subsection{Troubleshooting}
\label{subsec:trouble}
Here are some of the things that might go wrong on the initial
compilation and how to fix them.
\begin{itemize}
\item {\tt LaTeX Error: File `otagothesis.sty' not found}: \\
The simplest solution is to put the relevant
file(s) into the same directory as your document.  The files
\verb|otagothesis.sty| and \verb|logo.pdf| are in this distribution,
and \verb|natbib.sty|, \verb|plainnat.bst|, and \verb|moreverb.sty|
are all available on \verb|http://www.ctan.org|.
This can also happen if your latex distribution doesn't have
the \verb|natbib| package installed, in which case it will complain
about not having the files \verb|plainnat.bst| and \verb|natbib.sty|.
Other files that could be missing are \verb|logo.pdf| or
\verb|moreverb.sty|.  

\item The makefile doesn't work: \\
This makefile was written for GNUmake, and isn't guaranteed for any
other version of make.  If you need to compile without make, type the
following:
\begin{verbatim}
    pdflatex thesis
    bibtex thesis
    pdflatex thesis
    pdflatex thesis
\end{verbatim}
(also known as the \LaTeX\ stanza).  The extra {\tt pdflatex} at the end
is to get all the cross-references right.

\item I printed it up, and it's all off to one side!\\
It's supposed to be.  The bindery wants you to have a big binding
margin on the left of recto pages and on the right of verso pages.
If you are printing up using the one-sided option, then all of the
pages should have a wider margin on the left.
\
\item Any other problems are caused by one of two things: either you
have jumped the gun and edited a file, or your installation of \LaTeX\
doesn't look like mine.  You need the standard \LaTeXe\ packages for
it all to work; in particular the {\tt moreverb}, {\tt graphicx}, and
{\tt natbib} packages are absolutely crucial.

\item Many people in the department use \LaTeX , ask around for help! Internet is also great for finding solutions to \LaTeX\ problems.

\end{itemize}

\subsection{The Makefile}
\label{subsec:makefile}
The makefile for this package is called (appropriately enough)
``Makefile''.  You will find detailed instructions for its use in the
header comments.  A makefile tries to figure out what needs to be done
every time you compile your thesis, depending on what was last
changed.  It can save you a lot of typing.

Briefly: typing {\tt make} will create a pdf file viewable with
Acrobat reader, MacOsX's Preview, or Linux's gv.
Typing {\tt make clean} will remove all auxilliary
files, and {\tt make touch} will prepare for a complete re-compile.
If you want to suppress automatic viewing, find the \verb|AUTOVIEW|
variable in the makefile and set it to \verb|no|.

\section{The {\tt otagothesis} Package: Overview}
\label{sec:overview}

The {\tt otagothesis} package has been designed using the guidelines in
the booklet ``Notes on the Preparation of Theses'', available from the
reference desk at the Central Library.  I strongly suggest that you
obtain a copy of this and read it.

Open up the {\tt thesis.tex} file to see how the {\tt otagothesis}
package has been invoked and what sorts of options it gives you.  There
are lots of comments to explain what each part is for.

The package has the following features:
\begin{itemize}
\item A 30mm binding margin.  If you are printing up single sided,
this will be on the left of every page.  If you are printing up double
sided, it will be on the {\em left} of {\em recto} (right-hand-side)
pages, and on the {\em right} of {\em verso} (left-hand-side) pages;
\item Automatic title-page and contents page generation;
\item Automatic insertion of abstract and acknowledgements files;
\item A wide margins option for proofreading purposes (so your
supervisor has room for the red pen!);
\item Inclusion of the bibliography in the table of contents, renamed
``References''.
\end{itemize}
The following subsection lists all of the package's options in detail. 

\subsection{Package Options}
To use the package, all you need to do is include the line
\begin{quote}
\verb|\usepackage[option1,option2, ... ]{otagothesis}|
\end{quote}
in your main document.  With no options at all, this will give you
the commands available in the package (see the next subsection) and
will format your document with the correct margins.  It assumes you
are doing a PhD.
The available options are:
\begin{description}
\item[{\tt nofigures}] Suppresses the production of a ``list of figures''
page.
\item[{\tt notables}] Suppresses the production of a ``list of
tables'' page.
\item[\tt nolibrary] Suppresses the production of a library
  declaration page.
\item[{\em thesistype}] Enter a thesistype to produce the correct
title-page.  Available thesistypes are: {\tt phd, msc, ma, mcom,
interimsci, interimarts, interimcom, dipsci, diparts, dipcom, bschons,
bahons, bcomhons}.  The package will default to {\tt phd}.
\end{description}

\subsection{Package Commands}
There is only one command provided with the package, and it is
designed to help you print draft copies.  The command is
\begin{quote}
\verb|\frontstuff|
\end{quote}
and it places all the preface material before
Chapter~1.  If you don't issue this command in the document, you won't
get a title-page, table of contents, etc.---your thesis will just
print from Chapter~1.

% Physics
\section{Other packages included in the template}
Physics department has added some other useful package options.

\subsection{{\tt amssymb } and {\tt amsmath}}
These are the essential packages for writing math and physics formulas. 
For a short guide to usage in your latex document see 
\url{http://mirror.aut.ac.nz/CTAN/info/short-math-guide/short-math-guide.pdf}

For example (this is an example based on the above document):
\begin{quote}
\begin{verbatim}
\begin{equation}\label{first}
\alpha=b+c
\end{equation}
some intervening text
\begin{subequations}\label{grp}
\begin{align}
\alpha&=b+c\label{second}\\
d&=e+f+g\label{third}\\
h&=i+j\label{fourth}
\end{align}
\end{subequations}
\end{verbatim}
\end{quote}

Produces the following:

\begin{equation}\label{first}
\alpha=b+c
\end{equation}
some intervening text
\begin{subequations}\label{grp}
\begin{align}
\alpha &=b+c\label{second}\\
d&=e+f+g\label{third}\\
h&=i+j\label{fourth}
\end{align}
\end{subequations}


\subsection{{\tt siunitx}}
A great package for typing units in and out of equation mode! 
A comprehensive guide is available at 
\url{http://tug.ctan.org/macros/latex/exptl/siunitx/siunitx.pdf}. 
%
For example
\begin{quote}
\begin{verbatim}
\si{kg.m.s^{-1}} \\ 
\si{\kilogram\metre\per\second} \\ 
\si[per-mode=symbol]{\kilogram\metre\per\second} \\ 
\si[per-mode=symbol]{\kilogram\metre\per\ampere\per\second}
\end{verbatim}
\end{quote}
Produces:\\
%
\si{kg.m.s^{-1}} \\ \si{\kilogram\metre\per\second} \\ 
\si[per-mode=symbol]{\kilogram\metre\per\second} \\ 
\si[per-mode=symbol]{\kilogram\metre\per\ampere\per\second}


\verb|siunitx| is particularly useful when typing unit within an equations:
The standard equation mode will italicise letters:
\[ 1m + 1m = 2m\]
\verb|siunitx| prevents that:
\begin{quote}
\begin{verbatim}
\[ 1\si{.m} + 1\si{.m} = 2\si{.m} \]
\end{verbatim}
\end{quote}

\[ 1\si{.m} + 1\si{.m} = 2\si{.m}\]




\subsection{{\tt hyperref}}
The {\tt hyperref} package generates clickable links within the compiled pdf file.
Anytime you reference literature using \verb| \citet{} ,\citep{} | etc. 
or reference a Chapter, Figure or Table using \verb|\ref{} | 
\url{http://otago.ac.nz}




\section{Using the Thesis Template}
The main aim of this package is to allow you to write completely
vanilla \LaTeX\ and have everything come out in the right place.  All
you need to do is know how to create each chapter file, how to include
them in the main document, and how to print up a draft as well as the
main copy.

So go ahead and write chapters which have \LaTeX\ formatting commands
in them, and \verb|\include| them into the {\tt thesis.tex} file.
If you are using \verb|make|, add the name of the chapter file to the
makefile (including the \verb|.tex| suffix). Type {\tt make} and
bingo: one perfectly formatted thesis!

\subsection{Printing a Final Copy}
Make sure you have issued the command \verb|\frontstuff| directly
after \verb|\begin{document}|.  You should also comment out the line
beginning with \verb|\includeonly|, which restricts which chapters
you are going to print up.  If you want to include an abstract and
acknowledgements, you should have files {\tt abstract.tex} and {\tt
acknowledgements.tex} in the same directory as {\tt thesis.tex} and
they will be included automatically.

Now all you have to do is type {\tt make} to preview your document.
If you are using the {\tt twosided,openright} option,
there will be quite a few blank pages because new chapters will always
start on a recto (right-hand-side) page.

The final document is in PDF format, so you can print it either
through a pdf viewer such as Acrobat, or by using the \verb|pdftops|
program to produce a postscript version that you can send to the
printer with \verb|lpr|.

\subsection{Printing Specific Chapters}
Let's suppose that you have 4 chapters done and are working on
Chapter~5.  Compile time is getting pretty long; it would be nice if
you could generate only Chapter~5, but retain correct pagination and
references.  Here's how:
\begin{enumerate}
\item Comment out the \verb|\includeonly| line and type {\tt make} to
create a complete set of {\tt .aux} files.
\item Uncomment the \verb|\includeonly| line and enter the names of
the chapter{s} you wish to print up (without the {\tt .tex} extensions).
\item Type {\tt make} to preview the document.  Only the
desired chapters and their references should be printed up, but the
page numbers and cross-references should all be correct.
\end{enumerate}

\section{More to Come}
In Chapter~\ref{chap:bib} we will look at how to use the Harvard style
citations and references provided by the {\tt natbib} package.
Then, in Chapter~\ref{chap:diagrams}, we discuss how to include
figures in your document.
Chapter~\ref{chap:code} has some examples of how to include code
snippets in the body of your text, and how to include code in the
appendices. Next, in Chapter~\ref{chap:tables}, some examples of
\LaTeX\ tables are presented.  Finally, Chapter~\ref{chap:final}
gives some dos and don'ts regarding style and content.
